---
title: "Google Survey Analysis"
author: "Jeffrey Naecker"
date: "5/11/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, message=FALSE, echo=FALSE, warning = FALSE}
library(readxl)
library(ggplot2)
library(dplyr)
library(reshape2)
library(stringr)
library(tidyr)
library(hrbrthemes)

knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)

theme_set(theme_ipsum())
```

```{r import}
responses <- 
  read_excel("survey-rsa7c6dtoorshwj7dafuv3oas4.xls", sheet=3) %>%
  mutate(State = str_sub(Geography, -2, -1)) %>%
  mutate(`Response Time (ms)` = as.numeric(str_replace(`Response Time #1`, "ms", ""))) %>%
  select(-`Time (UTC)`, -`Publisher Category`, -`Parental Status`, -Weight, -Geography, -`Response Time #1`) %>%
  setNames(sub("Question #1 Answer( [0-9])*: ", "", names(.))) 

responses$`User ID` <- factor(responses$`User ID`)
responses$`Gender` <- factor(responses$`Gender`)
responses$`Age` <- factor(responses$`Age`)
responses$`State` <- factor(responses$`State`)
responses$`Urban Density` <- factor(responses$`Urban Density`, levels = c("Urban", "Suburban", "Rural", "Unknown"))
responses$Income <- factor(responses$Income, levels = c("$0-$24,999", "$25,000-$49,999", "$50,000-$74,999", "$75,000-$99,999", "$100,000-$149,999", "$150,000+", "I prefer not to say", "Unknown"))

items <- responses %>%
  melt(id.vars = c("User ID", "Gender", "Age", "State", "Urban Density", "Income", "Response Time (ms)"), variable.name = "Question", value.name = "Response") %>%
  mutate(Response = !is.na(Response), Question = factor(Question))

reasons <- items %>%
  filter(Response) %>%
  mutate(Reason = Question) %>%
  select(-Question, -Response)
```


## Descriptive Statistics

Below I plot histograms of the distribution of respondents by age, gender, income, and US state.

```{r}
ggplot(responses, aes(x = Age)) + stat_count() 
ggplot(responses, aes(x = Gender)) + stat_count()
ggplot(responses, aes(x = Income)) + stat_count() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggplot(responses, aes(x = State)) + stat_count() + theme(axis.text.x = element_text(size = 6))
```

## Reasons Not to Give More

This plot replicates the default view on Google's survey dashboard:

```{r}
items %>% 
  group_by(Question) %>% 
  summarize(Fraction = mean(Response)) %>%
  ggplot(aes(x = Question, y = Fraction)) +
  geom_bar(stat = "identity") + 
  coord_flip() + 
  scale_x_discrete(limits = rev(levels(reasons$Reason)))
```

We can break this up by gender to see that men and women's reason for not giving are actually pretty similar:

```{r}
items %>% 
  group_by(Question, Gender) %>% 
  summarize(Fraction = mean(Response)) %>%
  ggplot(aes(x = Question, y = Fraction, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() +
  scale_x_discrete(limits = rev(levels(reasons$Reason))) +
  scale_fill_manual(values=c("#E69F00", "#56B4E9", "#999999"))
```

Similarly, we can look at the excuse distribution by income level:

```{r}
items %>% 
  group_by(Question, Income) %>% 
  summarize(Fraction = mean(Response)) %>%
  ggplot(aes(x = Question, y = Fraction)) +
  geom_bar(stat = "identity") + 
  facet_wrap(~Income) +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(reasons$Reason)))
```

